Dynamic Transition Analysis with TIMES

### 2020-09-18 doc/2020-03-paper/discussion.tex
added figures
Former-commit-id: 397fc5a3eb6bd05763c89c592904ed924036023e
Former-commit-id: ed765d313eda47c8f62f47644dc28d64032fd6a0
adds clean results for syscost, net emissions
Former-commit-id: 5a0d7dd093d146c65fbf51a7f4887799ed403066
Former-commit-id: 8d484390baa9c26c0788b9cfa933ef94f0c1e0fc

>>>>

%Notes for Ansh:
%Importance of nuclear : cost of various transitions, flexibility by itself and when coupled with storage (Li-ion, H2). How it precludes the need for CCS.
%Urgency of transition : scenarios + sensitivity analysis.
%Li-ion storage: land use requirements for each case, qualitatively discuss Co, Ni requirements. Say Li is okay.
%AECs,PEMECs,SOECs need to be highly flexible with high availability factors.
%SOFCs: need for extremely responsive SOFCs with high availability factor, flexible operation (cite challenges from NREL).
%pemecs still part of cost optimal solutions
%identify key techs, same fcs needed always
%nuclear is key
%pws will have to be cheaper/cost competitive with electrolyzers to have a role
%main H2, role of CCS LNG - secondary alternative to H2 - as shown by correlation with SOFC inv cost. Not good on its own. Supplement to H2.
%H2 non trivial despite large range of perturbation
% FC flexibility, more storage/H2 less flex for CCS, nuclear

%urgency
As the results of all base case scenarios (Figs. \ref{scen1}-\ref{scen5}) indicate, rapid retirement of fossil fuels and deployment of carbon neutral technologies is urgently required to achieve emission reduction targets. All coal and oil plants must be shut down between 2025-2030, and natural gas use must be reduced dramatically. Japan must restart all of its existing reactors at full operating capacity by 2022. Since hydrogen plays a key role in zero to moderate nuclear deployment scenarios (Figs. \ref{scen3}and \ref{scen4}), it is prudent to invest in hydrogen power. As our model uses learning curves for technology prices and our objective function is the transition cost, our model deploys hydrogen technologies as late as possible in order to minimize  costs by utilizing technologies at their cheapest, while there is still time to deploy them and achieve emission goals. Therefore, our results indicate that it is imperative that Japan deploy an increasing amount of renewables including wind power between 2020-2030, and be prepared to deploy and rapidly scale utility-scale hydrogen by 2030 at the latest, in the absence of new nuclear, or by 2035, with new nuclear.

%nuclear
The scenario with the most nuclear deployment, Scenario 2, also has the lowest transition cost. When used with hydrogen, 50 MW of nuclear new nuclear results in the greatest reduction in emissions after 2050 out of all 5 scenarios. It also reduces the cost of the transition from 3,187,940 MUSD in Scenario 3 to 2,804,753 MUSD in Scenario 4. This is a difference of 383,187 MUSD. The estimated savings due to nuclear are conservative, as we have not incorporated the cost of transporting hydrogen through tankers or pipelines. Furthermore, the large emission reductions achieved with the help of nuclear allow natural gas to continue operating until 2100 while also meeting emission goals. As natural gas is excellent for peaking and load-following, when coupled with flexible fuel cell technology, it should result in an extremely stable energy system. Thus, in all scenarios, nuclear plays a key role in reducing emissions. Japan should strongly consider reinvigorating its nuclear energy sector by restarting existing plants, constructing new plants, and investing in advanced reactor research. To minimize the environmental impact from their life cycle emissions, it is vital that all existing and any new reactors be operated for a lifetime of 60 years or more at high capacity factors. Premature decommissioning due to operational problems or lack of public acceptance need to be avoided. Therefore, prioritizing reactor safety for resilience to disasters and in order to regain the public's trust in nuclear power, and increasing public awareness of nuclear's vital role in emission mitigation are key to achieving 2030 and 2050 emission targets.

%renewables and flexibility
In all scenarios, solar and wind supply a large part of the electricity demand. Unless large numbers of nuclear power plants are constructed, Japan must also invest in offshore wind farms by 2030 (Figs. \ref{scen1,scen3,scen4,scen5}). If hydrogen and CCS are not deployed by 2035, investment in offshore floating turbines may also necessary (\ref{scen1}). With such a large share of renewables, grid flexibility is extremely important. This requires significant investment in storage technologies, and ensuring that all emerging technologies are flexible and responsive. As natural gas is the only extant option for load-following, it is important to incorporate nuclear as in Scenarios 2 and 4 (Figs. \ref{scen2} and \ref{scen4}), or invest heavily in hydrogen (\ref{scen3} and \ref{scen5} to achieve emission cuts while simultaneously keeping natural gas operational for load-following. Additionally, nuclear power plants and CCS power plants must be able to load-follow, or be coupled to storage, ideally electrolysers, in order to store their excess electricity for peak demand. Any fuel cell technology utilized must also be extremely flexible and responsive. Since our model prefers \gls{SOFC}s to \gls{PEMFC}s due to their higher efficiency, it is important that \gls{SOFC}s reduce their startup times to have an edge over \gls{PEMFC}s in utility-scale applications.

%batteries
Lithium ion is the preferred storage medium in the absence of hydrogen. However, as the base case scenarios indicate, extremely large capacities of storage must be deployed to achieve a stable grid that can sustain the large capacities of renewables required for deep emission cuts. While lithium and cobalt may be available, cobalt and manganese reserves are limited \cite{scrosati_lithium-ion_2011,simon_potential_2015,turcheniuk_ten_2018} , which may inflate the prices of lithium ion storage if it is relied upon as the primary storage medium. It may be preferable to redesign batteries to reduce the amount of rare minerals used in their manufacturing, and improve recycling to increase the recovery rate of rare metals. The life cycle emissions from batteries must also be reduced by using cleaner materials and electricity for manufacturing. If hydrogen storage is available, battery storage storage serves as a near-term transition technology after which hydrogen storage dominates (\ref{scen3,scen4,scen5}).

%hydrogen - key players, key parameters, path forward?
After nuclear, hydrogen power emerges as the second most important technology for achieving 2050 emission goals. As seen in the sensitivity analysis results, it maintains a significant share of the electricity mix despite a wide range of perturbations to key parameters of the hydrogen sector. While \gls{SOFC}s and \gls{SOEC}s are preferred over \gls{PEMFC}s and \gls{PEMEC}s in all of our simulations, the role of hydrogen is so important that if \gls{SOFC} or \gls{SOEC} deployment is disabled, the model replaces them with inferior \gls{PEMFC}s or \gls{PEMEC}s respectively to recreate similar energy mixes. Key technologies that aid in decarbonization in our simulations are \gls{SOFC}s and \gls{PEMEC}s, which are close to maturation. Steam reforming, with or without CCS, does not get utilized in our simulations due to its high emission coefficient. At the lower end of the technology readiness scale, \gls{SOEC}s emerge as tremendously disruptive due to their high efficiency. However, their lifetime must be increased significantly, their life cycle emissions must be kept low, and they must become cost-competitive with \gls{PEMEC}s to realize their potential. \gls{PWS} plays a marginal role as it is not cost-competitive with electrolysis. From the simulation output data, the investment cost of \gls{SOFC}s emerges as an obviously important parameter. Therefore it is vital to reduce fuel cell investment costs to make deep emission reduction economically feasible. However, low response times and high availability factors are other key factors that are implicit in our assumptions. If \gls{SOFC}s are not as flexible as we have assumed, they are replaced in our simulations by \gls{PEMFC}s, which are known to be more flexible and are closer to commercialization. In Scenarios 3 and 4, solar, wind, and nuclear are used to generate hydrogen. Therefore it would be economically favourable to couple utility-scale renewables and nuclear power plants with electrolysis plants. Waste heat from nuclear could also be used to produce hydrogen, and would reduce reliance on renewables for electrolysis. As mentioned previously, more flexible and responsive fuel cells and electrolyzers are likely to be the dominant technology in such a system.

%ccs
CCS plays a small role in our base case scenarios as a transitional technology, despite being cheaper than hydrogen. Its share is found to be largely insensitive to its own investment cost in our sensitivity analysis. This is mainly due to its large emission coefficient. Marginal gains are expected in its penetration if the efficiency of CCS is increased to reduce its direct emissions. Reducing indirect emissions is likely to have a greater impact. This could be achieved in the future by increased electrification of the industrial sector, and the use of hydrogen to produce steel.

### 2020-09-18 doc/2020-03-paper/methods.tex
adds clean results for syscost, net emissions
Former-commit-id: 5a0d7dd093d146c65fbf51a7f4887799ed403066
Former-commit-id: 8d484390baa9c26c0788b9cfa933ef94f0c1e0fc

\begin{table}[!ht]
	\caption{Scenario definition.}
	\vspace{0.1in}
	\begin{tabularx}{\textwidth}{p{0.25\textwidth} p{0.25\textwidth} p{0.25\textwidth}p{0.25\textwidth}}
\hline 
\textbf{Technology} \textbf{Parameter} & \textbf{Sampled Distribution} & \textbf{Distribution Mean}& \textbf{Distribution Range}\\
\hline
\gls{PWS} Investment Cost       & Gaussian    & 3088 \$/kW & $\pm$20\% around the mean. \\                  
\gls{SOEC} Investment Cost      & Gaussian    & 1388 \$/kW & $\pm$20\% around the mean. \\                  
\gls{PEMEC} Investment Cost     & Gaussian    & 3800 \$/kW & $\pm$20\% around the mean. \\                  
\gls{SOFC} Investment Cost      & Gaussian    & 7399 \$/kW & $\pm$20\% around the mean. \\                  
\gls{PEMFC} Investment Cost     & Gaussian    & 7399 \$/kW & $\pm$20\% around the mean. \\                  
\gls{CCS} Gas Investment Cost   & Gaussian    & 2626 \$/kW & $\pm$20\% around the mean. \\                  
\gls{CCS} Coal Investment Cost  & Gaussian    & 5252 \$/kW & $\pm$20\% around the mean. \\
\gls{PWS} Emission Coefficient  & Triangular  & 1.08 g/kWh & 0.2-5.405 g/kWh. \\                  
\gls{SOEC} Emission Coefficient & Triangular  & 1.08 g/kWh & 0.2-5.405 g/kWh. \\
\gls{PWS} Efficiency            & Triangular  & 0.525 & 0.5-0.58. \\                               
\hline
	\end{tabularx}
\label{sa-vars}
\end{table}

>>>>

\begin{table}[!ht]
	\caption{Scenario definition.}
	\vspace{0.1in}
	\begin{tabularx}{1.1\textwidth}{p{0.3\textwidth} p{0.25\textwidth} p{0.25\textwidth}p{0.3\textwidth}}
\hline 
\textbf{Technology}  & \textbf{Sampled} & \textbf{Distribution}& \textbf{Distribution}\\
\textbf{Parameter} & \textbf{Distribution} & \textbf{Mean}& \textbf{Range}\\
\hline
\gls{PWS} Investment Cost       & Gaussian    & 3088 \$/kW & $\pm$20\% of mean. \\                  
\gls{SOEC} Investment Cost      & Gaussian    & 1388 \$/kW & $\pm$20\% of mean.\\                  
\gls{PEMEC} Investment Cost     & Gaussian    & 3800 \$/kW & $\pm$20\% of mean.\\                  
\gls{SOFC} Investment Cost      & Gaussian    & 7399 \$/kW & $\pm$20\% of mean.\\                  
\gls{PEMFC} Investment Cost     & Gaussian    & 7399 \$/kW & $\pm$20\% of mean.\\                  
\gls{CCS} Gas Investment Cost   & Gaussian    & 2626 \$/kW & $\pm$20\% of mean.\\                  
\gls{CCS} Coal Investment Cost  & Gaussian    & 5252 \$/kW & $\pm$20\% of mean.\\
\gls{PWS} Emission Coefficient  & Triangular  & 1.08 g/kWh & 0.2-5.405 g/kWh. \\                  
\gls{SOEC} Emission Coefficient & Triangular  & 1.08 g/kWh & 0.2-5.405 g/kWh. \\
\gls{PWS} Efficiency            & Triangular  & 0.525 & 0.5-0.58. \\                               
\hline
	\end{tabularx}
\label{sa-vars}
\end{table}

### 2020-09-18 doc/2020-03-paper/results.tex
adds clean results for syscost, net emissions
Former-commit-id: 5a0d7dd093d146c65fbf51a7f4887799ed403066
Former-commit-id: 8d484390baa9c26c0788b9cfa933ef94f0c1e0fc

The fifth scenario's results (Fig. \ref{scen5}) closely resemble those of the third scenario (Fig. \ref{scen3}). However, \gls{SOEC}s rapidly replace a large fraction of aging \gls{PEMEC}s. Some \gls{PEMEC}s remain in the mix until 2100 despite their lower efficiency due to their lower cost. Both 2030 and 2050 emission targets are met, but no additional emission reductions occur beyond 2050. The amount of electricity diverted to storage technologies is 41,752 TWh, primarily from solar and emerging solar technologies(48\%), onshore wind (28\%), and fixed-bottom offshore wind (20\%). From this, 35,550 TWh of hydrogen is generated, mostly from \gls{SOEC} (56\%), \gls{PEMEC} (43\%). \gls{AEC} contributes a mere 0.95\% of the total hydrogen, and \gls{PWS} plays the smallest role at 0.05\%. The cost of this transition is \textbf{3,180,613 MUSD}.

@@

\begin{figure}[h] 
\centering
\hspace*{-3cm}
\includegraphics[scale=0.14]{figures/co2emi}
\caption{Sensitivity analysis of the model's cumulative emissions over 2013-2100 with respect to selected model parameters. Red dots indicate points that are zero values, while blue dots indicate non-zero values. Plots with a grey background represent false correlations with characteristics of technologies that were not deployed (i.e. PEM FC and CCS Coal). }
\label{sa-co2}
\end{figure}

>>>>

The fifth scenario's results (Fig. \ref{scen5}) closely resemble those of the third scenario (Fig. \ref{scen3}). However, \gls{SOEC}s rapidly replace a large fraction of aging \gls{PEMEC}s, starting from 2055. Some \gls{PEMEC}s remain in the mix until 2100 despite their lower efficiency due to their lower cost. Both 2030 and 2050 emission targets are met, but no additional emission reductions occur beyond 2050. The amount of electricity diverted to storage technologies is 41,752 TWh, primarily from solar and emerging solar technologies(48\%), onshore wind (28\%), and fixed-bottom offshore wind (20\%). From this, 35,550 TWh of hydrogen is generated, mostly from \gls{SOEC} (56\%), \gls{PEMEC} (43\%). \gls{AEC} contributes a mere 0.95\% of the total hydrogen, and \gls{PWS} plays the smallest role at 0.05\%. The cost of this transition is \textbf{3,180,613 MUSD}.

@@

\begin{figure}[h] 
\centering
\hspace*{-3cm}
\includegraphics[scale=0.14]{figures/co2emi}
\caption{Sensitivity analysis of the model's cumulative emissions over 2013-2100 with respect to selected model parameters. Red dots indicate points that are zero values, while blue dots indicate non-zero values. Plots with a grey background represent false correlations with characteristics of technologies that were not deployed (i.e. PEM FC and CCS Coal).}
\label{sa-co2}
\end{figure}

As the first set of sensitivity analysis results (Fig. \ref{sa-techs}) indicate, \gls{PEMFC}s are never utilized due to their efficiency being lower than that of \gls{SOFC}s, and \gls{CCS} coal plants are never deployed due to their high emission coefficient. \gls{PWS} plays a minor role in generating 0-20 TWh of hydrogen over the entire simulation's time-period, as opposed to \gls{SOEC}s which generate 19,500-22,500 TWh of hydrogen, and \gls{PEMEC}s, which generate 12,000-16,000 TWh. The share of \gls{PWS} was found to be strongly correlated to its investment cost, with \$3300/kW the approximate threshold at which this technology ceases to be deployed. It's deployment is also weakly correlated with the splitting efficiency, wherein values greater than 21\% seem most favourable for its deployment. The emission coefficient is too small to have an impact on the overall emissions from the system considering \gls{PWS}'s small deployment, hence it does not appreciably affect its deployment. \gls{SOEC}'s share is most strongly correlated with their investment cost, however they play a non-trivial role over the range of perturbation applied to said parameter. Their deployment is largely insensitive to their emission coefficient as their cost and high efficiency trumps any increase in emissions due to their deployment, which is likely to be small due to their small emission coefficient. \gls{PEMEC} and \gls{SOFC} deployment is also most noticeably affected by their respective investment costs. The share of \gls{CCS}-Gas is relatively small, about 2-3\%, due to its emission coefficient being much larger than that of all hydrogen technologies. Its deployment is most strongly correlated with the investment cost of \gls{SOFC}s as a reduction in the deployment of fuel cells forces the model to marginally increase CCS Gas deployment to achieve its emission goals.

The second set of results (Fig. \ref{sa-syscost} and \ref{sa-co2}) indicate that system's total transition cost and total emissions are tightly correlated with the investment cost of \gls{SOFC}s. This follows from the large share of \gls{SOFC}s, and the fact that they are the most important technology in the model in terms of reducing emissions and converting excess power from renewables into electricity, yet they are the most expensive technology available to the model.
